Nanophotonics with optical phonons in two dimensions

### 2018-12-21 paper.tex
Nick's work from Overleaf

Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies. Phonon polaritons share many features in common with plasmon polaritons in conductors. In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra}. Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}. In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which shrink the in-plane and out-of-plane wavelength of polaritons monotonically with decreasing film thickness. In many cases, a monolayer is the ultimate limit of this effect. Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions. However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO-TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point.

In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions. Using these parameters for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected. We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced seven orders of magnitude by the emitter-LO phonon coupling, which due to the narrow bandwidth of phonon-polariton resonances, can lead to the possibility of realizing multi-mode strong coupling between an external emitter and optical phonons in hexagonal boron nitride.

@@

H_{\mathrm{int}} = \int d^2x \rho \phi = -\int d^2x (\nabla\cdot\mathbf{P})\phi,

@@

\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{\beta E_m}-\e^{\beta E_n} \right),

@@

where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, and $\beta$ is the inverse temperature. Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by: 

@@

where 'L' denotes longitudinal polarization, $|0_{\mathbf{q}L}\rangle$ ($|1_{\mathbf{q}L}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$. The frequency $\omega_{\mathbf{q}L}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO-TO splitting [Needs reference!]. This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility. Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response is an uncontrolled double-counting. The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  
To relate the polarization-polarization response function to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$. We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_1$ and a substrate of permittivity $\epsilon_2$.

@@

To proceed, we express the polarization matrix element in Equation (4) in terms of parameters of the monolayer such as the Born effective charges. Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in [Needs reference!], one finds that the RPA dielectric function is given by:
\begin{equation}
\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa,j} Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2}{\omega^2_{\mathrm{TO}}-\omega^2}
\end{equation}
In this expression, $\omega_L$ and been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO-TO splitting, they are degenerate. Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.

As is well known, the zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic waves. Thus, the condition for the zeros of the dielectric function can be expressed as 
\begin{align}
\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \nonumber \\
&= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
\end{align}
where $eZ^r = Z$, with $e$ the electron charge, and $V(q)$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates. This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon. In particular, it was shown in [Sohier2017] that in two-dimensional polar materials, the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO-TO splitting given by Equation (8) of the current manuscript. \textit{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.}

We now re-write $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit. From Equation (8), we may write Equation (7) as
\begin{equation}
\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{\omega^2_{\mathbf{q}\mathrm{LO}}-\omega^2_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2}\right).
\end{equation}
Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$. We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters. In many cases [ref], the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
\begin{equation}
\omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma \mathrm{TO}}+\mathcal{S}q} \approx \omega_{\Gamma \mathrm{TO}} + v_g q,
\end{equation}
where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_g$ is the group velocity of the LO mode, and $\omega_{\Gamma \mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point. The group velocity is specified in terms of microscopic parameters via 
\begin{equation}
v_g = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma TO}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
\end{equation}
Thus, in the long wavelength limit, we have the following universal parameterization of the RPA dielectric function of a polar monolayer:
\begin{equation}
\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{2v_g\omega_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\right),
\end{equation}
where we have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription. Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$. In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...

It has been well-appreciated in the literature on phonon polaritons in thin slabs. Here, we expand on the analogy in the two-dimensional case. To do so, we take Equation (8) in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \equiv  \left(\frac{eZ^r_*}{\sqrt{M_*}}\right)^2 \equiv \left(\frac{Q_*}{\sqrt{M_*}}\right)^2$. Then, the LO-TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$. Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$. This squared frequency can be thought of as the "LP-TP" splitting between longitudinal and transverse plasma oscillations. Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$". In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$", but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$. This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figure 1.

@@

In Fig. 2a, we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega) = -\frac{2\iu \omega}{\epsilon q}\left(\epsilon_{\mathrm{RPA}} - \epsilon_{\mathrm{env}} \right)$ below vacuum and on top of a substrate of permittivity $\epsilon_{\mathrm{s}}$, is given by
\begin{equation}
r_p(\mathbf{q},\omega) = \frac{(\epsilon_{\mathrm{s}}-1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{\mathrm{s}}+1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
\end{equation}
In Figs. 2b, we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$. The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, because as was shown in [Rivera], all interactions involving emission of electromagnetic energy by emitters, such as electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission, scale as some power of this confinement factor. Interestingly, in the case of a monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.  In particular, at frequencies of 1450 $cm^{-1}$, the LO phonon already has a confinement factor in excess of conf, corresponding to a phonon wavelength of 6896.55/conf nm, which is much shorter wavelength than any phonon polariton thus measured so far, as well as shorter wavelength than any plasmon that has been measured in the mid-IR, such as in graphene. In fact, this short wavelength is mostly out of the range of measurability in scattering near-field microscope (SNOM) measurements, which with current tip radii, are limited to measuring wavelengths in excess of 60 nm. This 2D LO phonon can be measured with SNOM at frequencies close to the TO phonon frequency, where the confinement is less, but this comes at the price of very high dissipation and correspondingly low propagation quality factor, as illustrated in Fig. 2c. This is simply because frequencies near the TO phonon, which is strongly absorbing, correspond to the highest dissipation (measured by $\text{Re }\sigma$ or $\text{Im }\epsilon_{\mathrm{RPA}}$). These considerations imply that access to the lower-loss and higher-confined portions of the phononic waves, in the absence of a sharp tip, requires either a free electron probe, as in EELS, or a quantum emitter probe, such as a molecule or quantum well [cite people].

@@

\section{Strong light-matter interactions enabled by 2D optical phonons}

The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields. Applications of these couplings can be to realize ultra-bright single- or two-photon sources, or realize strong-coupling and the associated phenomenology of Rabi oscillations and polaritons, or resolve spectroscopically "forbidden" transitions. In Figure 3, we show key figures of emitter-phonon coupling in polar monolayers.

In Fig. 3a, we show the enhancement (or Purcell factor) $F_p$ of spontaneous emission of 2D LO phonons at frequency $\omega_0$ by an external emitter at different distances $d$ away from an infinite sheet of two-dimensional boron nitride. The rate of spontaneous emission $\Gamma$ into 2D LO phonons obeys
\begin{equation}
\Gamma = \frac{3c^3}{2\omega_0^3}\int dq~q^2e^{-2qz}\text{Im }r_p(q,\omega_0) \times \Gamma_0 \equiv F_p(\omega_0)\Gamma_0,
\end{equation}
where $\Gamma_0$ is the free-space spontaneous emission rate. As can be seen from Fig. 3a, for an emitter position 5 nm away from the surface, the Purcell factor can be very high, corresponding to over a million-fold increase in the rate of 2D LO phonon emission compared to photon emission into free-space. While this rate-enhancement is already quite high, it is a natural question how this coupling can be improved by nanostructuring, which will take the local density of electromagnetic states of Fig. 3a and concentrate it around frequencies associated with resonances of the nanostructure.For simplicity, we consider the effect of nanostructuring the hBN into a disk, which leads to the formation of modes as in Fig. 3b, in which we plot the surface charge density associated with each resonant mode. In Fig. 3c, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed 5 nm away from the surface of the disk. We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field. For an infrared emitter at a transition wavelength of 7 $\mu$m with a free-space radiative lifetime of 1 $\mu$s, 5 nm away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $2 \times 10^{14}$ $ s^{-1}$). This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling. Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultrastrong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons.

>>>>

	Phonon polaritons, hybrid quasiparticles of photons and optical phonons supported in polar materials, offer great promise for deeply sub-diffractional control of electromagnetic fields at mid-IR and THz frequencies. Phonon polaritons share many features in common with plasmon polaritons in conductors. In recent years, it has been shown that phonon-polaritons enable confinement of light to volumes over $10^6$ times smaller than that of a diffraction-limited photon in free-space\cite{caldwell2013low,xu2014mid,caldwell2014sub,dai2014tunable,tomadin2015accessing,yoxall2015direct,li2015hyperbolic,dai2015subdiffractional,dai2015graphene,caldwell2015low,li2016reversible,Basov:2016,basov2017towards,low2017polaritons,giles2017ultra,li2018infrared,ma2018plane}. Due to this remarkable confinement and their relatively high lifetimes of around picoseconds, phonon-polaritons open new opportunities for vibrational spectroscopy, radiative heat transfer \cite{hillenbrand2002phonon}, and control of dynamics in quantum emitters \cite{kumar2015tunable,rivera2017making,kurman2018control}. In many cases, extreme confinement of phonon polaritons is achieved by the use of thin-films, which shrink the in-plane and out-of-plane wavelength of polaritons monotonically with decreasing film thickness \cite{dai2014tunable,dubrovkin2018ultra}. In many cases, a monolayer is the ultimate limit of this effect. Thus, to bring these classical and quantum nanophotonics applications to their ultimate limits, it is necessary to consider the properties of these modes in two-dimensions. However, this picture of monolayers supporting maximally confined phonon-polariton modes is ostensibly complicated by the fact that in two-dimensional polar materials, the LO-TO splitting that gives rise to phonon-polaritons in three-dimensions is absent at the $\Gamma$ point \cite{sanchez2002vibrational,mele2002electric,serrano2007vibrational,sohier2017breakdown}.

	In this Letter, we develop core results in the nascent field of optical phononics in two-dimensional materials by deriving a universal form for the dielectric function of a polar monolayer, incorporating spatially nonlocal effects, which is specified in fully in terms of the dispersion relation of LO and TO phonons in two dimensions. Using parameters from Ref. \cite{sohier2017breakdown} for important monolayers such as hexagonal boron nitride, we present the confinement and propagation losses of these modes, identifying the frequency region where they should be most easily detected. We then show that these modes essentially provide the ultimate limit of light-matter interaction between emitters and polar materials, showing that for atom-like emitters, their spontaneous decay can be enhanced by up to eight orders of magnitude by the emitter-LO phonon coupling, which due to the narrow bandwidth of phonon-polariton resonances, can lead to the possibility of realizing the multi-mode strong coupling and ultra-strong coupling regimes of quantum electrodynamics for an external emitter coupled with optical phonons in hexagonal boron nitride.

@@

	H_{\mathrm{int}} = \int d^2x ~\rho \phi = -\int d^2x~ (\nabla\cdot\mathbf{P})\phi,

@@

	\boldsymbol{\Pi}(\mathbf{q},\omega) =  \frac{1}{\epsilon_0 \Omega\mathcal{Z}}\sum\limits_{m,n}\frac{\mathbf{P}_{mn}(\mathbf{q})\otimes\mathbf{P}_{nm}(\mathbf{q})}{\hbar\omega + E_{nm}+\iu 0^+}\left(\e^{-\beta E_m}-\e^{-\beta E_n} \right),

@@

	where $m,n$ are states of the phononic Fock space of the monolayer, $P_{i,mn} = \sum\limits_{\kappa,j}Z_{\kappa,ij}\langle m | u_{\kappa,j} | n \rangle$ are matrix elements of the polarization associated with phonon modes, $E_{m (n)}$ is the energy of state $m$ $(n)$, $\beta$ is the inverse temperature, and $\mathcal{Z}$ is the grand partition function. Evaluating the contribution of optical phonons in the harmonic approximation to the polarization-polarization response, one finds that the response function is given by: 

@@

	where 'L' denotes longitudinal polarization, $|0_{\mathbf{q}L}\rangle$ ($|1_{\mathbf{q}L}\rangle$) denotes a state with no (one) longitudinal phonon of wavevector $\mathbf{q}$. The frequency $\omega_{\mathbf{q}L}$ in the denominator, as in the case of bulk phonons, is the frequency of the longitudinal phonon of wavevector $\mathbf{q}$ prior to considering LO-TO splitting [Needs reference!]. This is the consistent with the fact that LO-TO splitting is a collective effect arising from Coulomb interactions and the fact that the equation above represents a single-particle susceptibility. Coulomb interactions are accounted for in the random phase approximation, and to include them in the single-particle response is an uncontrolled double-counting. The component $qq$ in the response tensor denotes a pair of directions parallel to the wavevector.  To relate the polarization-polarization response function to the electromagnetic modes supported by a polar monolayer, we solve Maxwell's equations for an evanescent electromagnetic mode supported by a surface with polarization-polarization response tensor $\mathbf{\Pi}$. We consider the monolayer to be sandwiched by a superstrate of permittivity $\epsilon_1$ and a substrate of permittivity $\epsilon_2$.

@@

	To proceed, we express the polarization matrix element in Equation (4) in terms of parameters of the monolayer such as the Born effective charges. Considering the longitudinal phonon contribution to the second-quantized ionic displacement, as in \cite{srivastava1990physics}, one finds that the RPA dielectric function is given by:
	\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}} - \frac{q}{2\epsilon_0}\frac{\Big|\hat{q}\cdot\sum\limits_{\kappa,j} Z_{\kappa,ij}\eta_{\kappa,j}  \Big|^2}{\omega^2_{\mathrm{TO}}-\omega^2}
	In this expression, $\omega_L$ has been re-named as $\omega_{\mathrm{TO}}$, because in the absence of LO-TO splitting, they are degenerate. Additionally, we have defined eigendisplacements $\boldsymbol{\eta}_{\kappa\mathbf{q}}\equiv \frac{e_{\kappa\mathbf{q},L}}{\sqrt{M_{\kappa}}}$, where $e_{\kappa\mathbf{q},L}$ is the unit-normalized polarization vector of atom $\kappa$ in the unit cell oscillating according to a longitudinal phonon of wavevector $\mathbf{q}$ and $M_{\kappa}$ is the mass of atom $\kappa$.
	\begin{figure}[t]
		\includegraphics[width=85mm]{fig1.pdf}
		\caption{\textbf{LO phonons as the basic electromagnetic waves of a polar monolayer.} (a) Schematic structure of a polar monolayer such as hexagonal boron nitride. (b) Properties of LO and TO phonons in 3D and 2D. In 3D, there is a finite LO-TO splitting at zero wavevector, while in 2D there is none. Despite this, the 2D LO phonon plays the role of the phonon polariton in 3D and thin films. (c) Analogous physics appears in electron gases in 3D and 2D, where the 3D plasma frequency is analogous to the 3D LO-TO splitting. In 2D, the plasma frequency at zero wavevector is zero, but the electromagnetic physics is determined by the dispersion of 2D plasmons, which replace the plasmon polariton of bulk and thin films.}
	\end{figure}
	As is well known, the zeros of the RPA dielectric function, denoted $\omega_{\mathbf{q}}$, correspond to longitudinal electromagnetic waves. Thus, the condition for the zeros of the dielectric function can be expressed as 
	\begin{align}
	\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} &= \frac{e^2}{2\epsilon_0\epsilon_{\mathrm{env}}q}\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \nonumber \\
	&= V(q)\frac{1}{\Omega}q^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2,
	\end{align}
	where $eZ^r = Z$, with $e$ the electron charge, and $V(q)$ is the Coulomb interaction in two dimensions, screened by the dielectric environment of the sub- and superstrates. This result, as we now show, reveals that the TM mode corresponds simply to the two-dimensional optical phonon. As shown recently in \cite{sohier2017breakdown}, in two-dimensional polar materials the extra restoring forces on LO phonons relative to TO phonons, due to the Coulomb interaction, lead to a wavevector-dependent LO-TO splitting. The expression for the LO-TO splitting is none other than Equation (8) of the current manuscript. \textit{Thus, one of the main results of our manuscript is that the wave solutions Maxwell's equations associated with phonon-polaritons in bulk materials, correspond to pure LO phonons in the monolayer limit.} 
	\begin{figure*}[t]
		\includegraphics[width=135mm]{fig2.pdf}
		\caption{\textbf{Propagation characteristics of 2D LO phonons in an un-structured monolayer of boron nitride.} (a) Imaginary part of the Fresnel reflection coefficient for p-polarized waves in free-standing monolayer hBN, which peaks at the dispersion relation of the 2D LO phonon. (b) Confinement factor of 2D LO phonon modes, which measures how much smaller the 2D LO phonon wavelength is compared to that of a photon at the same frequency. (c) Propagation quality factor, which measures the number of wavelengths of propagation of the 2D LO phonon mode, for different values of the damping rate of phonon.}
	\end{figure*}
	We now re-write $\epsilon_{\mathrm{RPA}}$ in a way that is expressed purely in terms of the optical phonon dispersion, derive the conductivity of the monolayer, and derive a universal form for the dispersion of 2D LO phonons in the long-wavelength limit. From Equation (8), we may write Equation (7) as
	\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{\omega^2_{\mathbf{q}\mathrm{LO}}-\omega^2_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2}\right).
	Similarly, the conductivity is prescribed by $\sigma(\mathbf{q},\omega) = -\iu \omega \Pi(\mathbf{q},\omega)$. We are now in a position to derive a universal form for the RPA dielectric function of a 2D polar slab in terms of three phenomenological parameters. In many cases [ref], the Born charges are independent of $q$ for small $q$, and thus the dispersion can be written as
	\omega_{\mathbf{q},\mathrm{LO}} = \sqrt{\omega^2_{\Gamma \mathrm{TO}}+\mathcal{S}q} \approx \omega_{\Gamma \mathrm{TO}} + v_g q,
	where $\mathcal{S}$ is a microscopic constant determined by the Born charges of the monolayer, $v_g$ is the group velocity of the LO mode, and $\omega_{\Gamma \mathrm{TO}}$ is the TO phonon frequency at the $\Gamma$ point. The group velocity is specified in terms of microscopic parameters via 
	v_g = \frac{e^2 }{4\epsilon_0 \epsilon_{\mathrm{env}}\omega_{\Gamma TO}\Omega}\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2. 
	Thus, in the long wavelength limit, we have the following universal parameterization of the RPA dielectric function of a polar monolayer:
	\epsilon_{\mathrm{RPA}}(\omega) = \epsilon_{\mathrm{env}}\left( 1 + \frac{2v_g\omega_{\mathbf{q}\mathrm{TO}}}{\omega^2_{\mathbf{q}\mathrm{TO}}-\omega^2+\iu \omega \tau^{-1}}\right),
	where we have also phenomenologically included the phonon dissipation rate $\tau^{-1}$, consistently with  a relaxation-time prescription. Thus, all of the optical properties associated with LO phonons is captured by the TO phonon frequency, the slope of the LO dispersion at $\Gamma$, and the dissipation $\tau^{-1}$. %In Table 1, we tabulate these parameters for a few polar monolayers whose phonon band-structures have been calculated in the literature: hBN, MoS$_2$, and ...

	It has been well-appreciated in the literature on phonon polaritons in thin slabs. Here, we expand on the analogy in the two-dimensional case. To do so, we take Equation (8) in the case of a two-atom unit cell (such as hBN), and note that the term in the sum over Born charges can be written as $e^2\Big|\sum\limits_{\kappa,j}Z^r_{\kappa,ij}\eta_{\kappa,j}  \Big|^2 \equiv  \left(\frac{eZ^r_*}{\sqrt{M_*}}\right)^2 \equiv \left(\frac{Q_*}{\sqrt{M_*}}\right)^2$. Then, the LO-TO splitting can be written as $\omega^2_{\mathbf{q}} - \omega^2_{\mathbf{q},\mathrm{TO}} = \frac{Q_*^2}{2\epsilon_0\epsilon_{\mathrm{env}} M_*}q$. Now we note that the RHS is exactly the squared-frequency $\omega^2_{\mathbf{q}p}$ for a plasma oscillation in a 2D gas of charged particles with charge $Q_*$ and mass $M_*$. This squared frequency can be thought of as the "LP-TP" splitting between longitudinal and transverse plasma oscillations. Of course, there are no transverse plasma oscillations, and so ``$\omega_{\mathrm{TP}} = 0$". In the three-dimensional plasmon case, `$\omega_{\mathrm{TP}} = 0$", but the difference between the squared longitudinal and transverse plasma oscillation frequencies at zero-wavevector is non-zero and given by $\omega_p^2$. This analogy between phononic and plasmonic behavior as a function of dimension is illustrated schematically in Figures 1(b,c).

@@

	In Fig. 2a, we present the dispersion relation, operationally defined through the poles of the Fresnel reflection coefficient for $p$-polarized waves, which for a monolayer of conductivity $\sigma(\mathbf{q},\omega) = -\frac{2\iu \omega}{\epsilon q}\left(\epsilon_{\mathrm{RPA}} - \epsilon_{\mathrm{env}} \right)$ below vacuum and on top of a substrate of permittivity $\epsilon_{\mathrm{s}}$, is given by
	\begin{equation}
	r_p(\mathbf{q},\omega) = \frac{(\epsilon_{\mathrm{s}}-1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega\epsilon_0}}{(\epsilon_{\mathrm{s}}+1)-\frac{\sigma(\mathbf{q},\omega) q}{\omega \epsilon_0}}.
	\end{equation}
	In Figure 2b, we show the confinement factor (or effective mode index) of the 2D LO phonon mode, defined by $n_{\mathrm{eff}} = \frac{qc}{\omega}$. The confinement factor is a key figure of merit in many nanophotonic applications involving quantum emitters, because as was shown in \cite{archambault2010quantum,koppens2011graphene,rivera2016shrinking,rivera2017making}, all interactions involving emission of electromagnetic energy by emitters, such as electric and magnetic dipole emission, multipole emission, phosphorescence, and multiphoton emission, scale as some power of this confinement factor. Interestingly, in the case of a monolayer, the confinement grows very rapidly in frequency, due to the extremely low group velocity of 2D LO phonons, which is a remarkable four orders of magnitude slower than the speed of light.  In particular, at frequencies of 1450 $cm^{-1}$, the LO phonon already has a confinement factor in excess of 400, corresponding to a phonon wavelength below 20 nm, which is much shorter wavelength than any phonon polariton thus measured so far, as well as shorter wavelength than any plasmon that has been measured in the mid-IR, such as in graphene. In fact, this short wavelength is mostly out of the range of measurability in current scattering near-field microscope (SNOM) measurements, which with current tip radii, are limited to measuring wavelengths no shorter than 60 nm. This 2D LO phonon can be measured with SNOM at frequencies close to the TO phonon frequency, where the confinement is less, but this comes at the price of very high dissipation and correspondingly low propagation quality factor, as illustrated in Fig. 2c, where Re $q$/Im $q$ < 5. This is simply because frequencies near the TO phonon resonance correspond to very high dissipation (measured by $\text{Re }\sigma$ or $\text{Im }\epsilon_{\mathrm{RPA}}$). These considerations imply that access to the lower-loss and higher-confined portions of the phononic waves, in the absence of a sharp tip, requires either a free electron probe, as in EELS, where slow electrons can be used to probe plasmon wavelengths of just a few nanometers in monolayer metals, as well as the nonlocal bulk plasmon dispersion in metals \cite{nagao2001dispersion,diaconescu2007acoustic,de2010optical}. EELS has been recently employed to measure phonon polaritons in ultrathin films of hBN \cite{govyadinov2017probing}. Yet another interesting type of probe, with relevance to fundamental physics and quantum optics applications, is a quantum emitter such as an atom, molecule, or quantum well. A quantum emitter, which one can consider as a sub-nanoscale tip, can see very high wavevectors of the 2D LO phonon if sufficiently close to the polar monolayer. Recently, it was demonstrated using nanostructures of bulk hBN that the interaction of vibrational emitters with phonon polaritons is on the border of the strong coupling regime \cite{autore2018boron}.

@@

	\section{Strong light-matter interactions enabled by 2D optical phonons}
	\begin{figure}[t]
		\includegraphics[width=80mm]{fig3.pdf}
		\caption{\textbf{Extreme spontaneous emission enhancement due to 2D LO phonons in nanostructured geometries.} Enhancement of the spontaneous emission rate for an emitter concentric with the disk at a distance $z=5$ nm away from the disk. For a disk with a diameter of 20 nm, the rate of emission enhancement can be 100 million-fold. For an emitter with a free-space decay rate of $10^6$ cm$^{-1}$ at 7 $\mu$m, the emitter would experience a decay rate comparable to the frequency of the disk mode, leading to ultra-strong coupling of an external emitter with 2D LO phonons.}
	\end{figure}
	The extreme confinement of electromagnetic fields offered by the 2D LO phonon presents an opportunity for quantum optical applications in which one seeks to couple an external emitter such as an atom, molecule, defect, or artificial atomic system to electromagnetic fields. Applications of these couplings can be to realize ultra-bright single- or two-photon sources, or realize strong-coupling and the associated phenomenology of Rabi oscillations and polaritons, or resolve spectroscopically "forbidden" transitions.

	In Fig. 3, we consider the quantum coupling of a dipole emitter to 2D LO phonon modes in nanostructured monolayer hBN. For simplicity, we consider hBN nanostructured as a disk, which leads to the formation of sharp resonances. In Fig. 3, we plot the enhancement of the spontaneous emission rate for an external emitter polarized perpendicularly to the plane of the disk and placed 5 nm away from the surface of the disk. We find that the rate of spontaneous emission of 2D optical phonons is approximately 8 orders of magnitude larger than the rate of spontaneous emission of photons in the far field. For an infrared emitter at a transition wavelength of 7 $\mu$m with a free-space radiative lifetime of 1 $\mu$s, 5 nm away from an hBN disk, the coupling rate to 2D optical phonons would be on the same scale as the optical phonon frequency itself (about $2 \times 10^{14}$ $ s^{-1}$). This coupling rate thus implies coupling between an emitter and the field in the regime of ultra-strong coupling. Thus, the extreme confinement of electromagnetic energy associated with LO phonons in two dimensions enables the possibility of realizing ultrastrong coupling of an atom or molecule with optical phonons in a polar material, allowing the potential realization of new coupled states of quantum emitters and phonons.
